---
title: "Microgastropods as Environmental Indicators"
author:
  - name: "Kelmo et al."
    affiliation: "GEEMCO Lab"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    code_folding: hide
  pdf_document:
    toc: true
    number_sections: true
bibliography: references.bib
csl: ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  dpi = 300
)

# Load required packages
library(tidyverse)
library(here)
library(knitr)
library(patchwork)
```

# Abstract

**Background:** *Provide context about microgastropods and their potential as environmental indicators.*

**Methods:** *Describe sampling methods, study sites, and analytical approaches.*

**Results:** *Summarize key findings about diversity patterns and environmental relationships.*

**Conclusions:** *State main conclusions about microgastropods as environmental indicators.*

**Keywords:** microgastropods, environmental indicators, biodiversity, freshwater ecology, community ecology

# Introduction

## Background

Microgastropods represent an important but often overlooked component of aquatic ecosystems. Their sensitivity to environmental conditions makes them potentially valuable as bioindicators.

## Research Questions

1. How do microgastropod communities vary across environmental gradients?
2. Which environmental variables best predict microgastropod diversity?
3. Can specific species serve as indicators of environmental conditions?

## Objectives

- Characterize microgastropod community composition across study sites
- Analyze relationships between diversity and environmental variables
- Identify potential indicator species

# Methods

## Study Area

*Describe the geographic location, climate, and habitat characteristics of study sites.*

## Sampling Design

*Detail the sampling methodology, including:*
- Number of sites sampled
- Sampling dates and frequency
- Collection methods
- Sample size per site

## Environmental Variables

*List and describe measured environmental variables:*
- Water quality parameters (pH, temperature, dissolved oxygen, conductivity)
- Substrate characteristics
- Vegetation coverage
- Other relevant factors

## Laboratory Methods

*Describe specimen processing and identification procedures.*

## Data Analysis

```{r load-data, eval=FALSE}
# Load processed data
diversity_data <- read_csv(here("data", "processed", "diversity_with_metadata.csv"))
community_matrix <- read_csv(here("data", "processed", "community_matrix.csv"))
environmental_matrix <- read_csv(here("data", "processed", "environmental_matrix.csv"))
```

### Diversity Indices

Species richness, Shannon diversity, Simpson diversity, and evenness were calculated using the `vegan` package.

### Ordination Analysis

Non-metric multidimensional scaling (NMDS) was used to visualize community composition patterns.

### Statistical Tests

- Correlation analyses between diversity indices and environmental variables
- Environmental fitting to ordination results
- Indicator species analysis

# Results

## Community Composition

```{r community-overview, eval=FALSE}
# Summary statistics
summary_stats <- diversity_data %>%
  summarise(
    n_sites = n_distinct(site_id),
    mean_richness = mean(richness, na.rm = TRUE),
    sd_richness = sd(richness, na.rm = TRUE),
    mean_shannon = mean(shannon, na.rm = TRUE),
    sd_shannon = sd(shannon, na.rm = TRUE)
  )

kable(summary_stats, caption = "Summary of diversity indices across all sites")
```

## Diversity Patterns

### Species Richness

```{r richness-plot, eval=FALSE, fig.cap="Distribution of species richness across sampling sites"}
ggplot(diversity_data, aes(x = site_id, y = richness)) +
  geom_boxplot(fill = "steelblue", alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(
    title = "Species Richness by Site",
    x = "Site",
    y = "Species Richness"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Shannon Diversity

*Present results for Shannon diversity patterns.*

## Environmental Relationships

```{r diversity-environment, eval=FALSE, fig.cap="Relationships between diversity and environmental variables"}
# Example: Shannon diversity vs temperature
ggplot(diversity_data, aes(x = temperature, y = shannon)) +
  geom_point(size = 3, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Shannon Diversity vs Temperature",
    x = "Temperature (°C)",
    y = "Shannon Diversity Index"
  ) +
  theme_bw()
```

## Ordination Results

```{r ordination, eval=FALSE, fig.cap="NMDS ordination of microgastropod communities"}
# NMDS plot would be generated here
# Include interpretation of ordination axes and environmental vectors
```

## Indicator Species

*Present results of indicator species analysis, identifying species associated with specific environmental conditions.*

# Discussion

## Main Findings

*Interpret the key results in the context of existing literature.*

## Microgastropods as Environmental Indicators

*Discuss the implications for using microgastropods as bioindicators:*
- Sensitivity to environmental changes
- Practical considerations for monitoring
- Advantages and limitations

## Comparison with Other Studies

*Compare findings with previous research on microgastropods or similar taxa.*

## Limitations

*Acknowledge limitations of the study:*
- Sampling constraints
- Geographic scope
- Temporal resolution
- Other considerations

## Future Directions

*Suggest avenues for future research.*

# Conclusions

*Provide clear, concise conclusions addressing the research questions.*

# Acknowledgments

*Acknowledge funding sources, field assistants, and other contributors.*

# References

*References will be automatically generated from the bibliography file.*

# Supplementary Materials

## Supplementary Tables

```{r supplementary-tables, eval=FALSE}
# Include additional tables if needed
```

## Supplementary Figures

```{r supplementary-figures, eval=FALSE}
# Include additional figures if needed
```

---

## Session Information

```{r session-info}
sessionInfo()
```
