---
title: "Microgastropods as Environmental Indicators"
author:
  - name: "Kelmo et al."
    affiliation: "GEEMCO Lab"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: cosmo
    embed-resources: true
  pdf:
    toc: true
    number-sections: true
    documentclass: article
  docx:
    toc: true
    number-sections: true
bibliography: references.bib
csl: ecology.csl
execute:
  echo: true
  warning: false
  message: false
---

## Abstract

**Background:** *Provide context about microgastropods and their potential as environmental indicators.*

**Methods:** *Describe sampling methods, study sites, and analytical approaches.*

**Results:** *Summarize key findings about diversity patterns and environmental relationships.*

**Conclusions:** *State main conclusions about microgastropods as environmental indicators.*

**Keywords:** microgastropods, environmental indicators, biodiversity, freshwater ecology, community ecology

## Introduction

### Background

Microgastropods represent an important but often overlooked component of aquatic ecosystems. Their sensitivity to environmental conditions makes them potentially valuable as bioindicators.

### Research Questions

1. How do microgastropod communities vary across environmental gradients?
2. Which environmental variables best predict microgastropod diversity?
3. Can specific species serve as indicators of environmental conditions?

### Objectives

- Characterize microgastropod community composition across study sites
- Analyze relationships between diversity and environmental variables
- Identify potential indicator species

## Methods

### Study Area

*Describe the geographic location, climate, and habitat characteristics of study sites.*

### Sampling Design

*Detail the sampling methodology, including:*

- Number of sites sampled
- Sampling dates and frequency
- Collection methods
- Sample size per site

### Environmental Variables

*List and describe measured environmental variables:*

- Water quality parameters (pH, temperature, dissolved oxygen, conductivity)
- Substrate characteristics
- Vegetation coverage
- Other relevant factors

### Laboratory Methods

*Describe specimen processing and identification procedures.*

### Data Analysis

```{r}
#| label: setup
#| include: false

# Load required packages
library(tidyverse)
library(here)
library(knitr)
library(patchwork)
library(vegan)
```

```{r}
#| label: load-data
#| eval: false

# Load processed data
diversity_data <- read_csv(here("data", "processed", "diversity_with_metadata.csv"))
community_matrix <- read_csv(here("data", "processed", "community_matrix.csv"))
environmental_matrix <- read_csv(here("data", "processed", "environmental_matrix.csv"))
```

#### Diversity Indices

Species richness, Shannon diversity, Simpson diversity, and evenness were calculated using the `vegan` package.

#### Ordination Analysis

Non-metric multidimensional scaling (NMDS) was used to visualize community composition patterns.

#### Statistical Tests

- Correlation analyses between diversity indices and environmental variables
- Environmental fitting to ordination results
- Indicator species analysis

## Results

### Community Composition

```{r}
#| label: community-overview
#| eval: false
#| tbl-cap: "Summary of diversity indices across all sites"

# Summary statistics
summary_stats <- diversity_data %>%
  summarise(
    n_sites = n_distinct(site_id),
    mean_richness = mean(richness, na.rm = TRUE),
    sd_richness = sd(richness, na.rm = TRUE),
    mean_shannon = mean(shannon, na.rm = TRUE),
    sd_shannon = sd(shannon, na.rm = TRUE)
  )

kable(summary_stats)
```

### Diversity Patterns

#### Species Richness

```{r}
#| label: fig-richness
#| eval: false
#| fig-cap: "Distribution of species richness across sampling sites"
#| fig-width: 8
#| fig-height: 6

ggplot(diversity_data, aes(x = site_id, y = richness)) +
  geom_boxplot(fill = "steelblue", alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(
    title = "Species Richness by Site",
    x = "Site",
    y = "Species Richness"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Shannon Diversity

*Present results for Shannon diversity patterns.*

### Environmental Relationships

```{r}
#| label: fig-diversity-environment
#| eval: false
#| fig-cap: "Relationships between diversity and environmental variables"
#| fig-width: 8
#| fig-height: 6

# Example: Shannon diversity vs temperature
ggplot(diversity_data, aes(x = temperature, y = shannon)) +
  geom_point(size = 3, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Shannon Diversity vs Temperature",
    x = "Temperature (°C)",
    y = "Shannon Diversity Index"
  ) +
  theme_bw()
```

### Ordination Results

```{r}
#| label: fig-ordination
#| eval: false
#| fig-cap: "NMDS ordination of microgastropod communities"
#| fig-width: 8
#| fig-height: 6

# NMDS plot would be generated here
# Include interpretation of ordination axes and environmental vectors
```

### Indicator Species

*Present results of indicator species analysis, identifying species associated with specific environmental conditions.*

## Discussion

### Main Findings

*Interpret the key results in the context of existing literature.*

### Microgastropods as Environmental Indicators

*Discuss the implications for using microgastropods as bioindicators:*

- Sensitivity to environmental changes
- Practical considerations for monitoring
- Advantages and limitations

### Comparison with Other Studies

*Compare findings with previous research on microgastropods or similar taxa.*

### Limitations

*Acknowledge limitations of the study:*

- Sampling constraints
- Geographic scope
- Temporal resolution
- Other considerations

### Future Directions

*Suggest avenues for future research.*

## Conclusions

*Provide clear, concise conclusions addressing the research questions.*

## Acknowledgments

*Acknowledge funding sources, field assistants, and other contributors.*

## References

::: {#refs}
:::

## Supplementary Materials

### Supplementary Tables

```{r}
#| label: supplementary-tables
#| eval: false

# Include additional tables if needed
```

### Supplementary Figures

```{r}
#| label: supplementary-figures
#| eval: false

# Include additional figures if needed
```

## Session Information

```{r}
#| label: session-info

sessionInfo()
```
